Supplementary Materials And Methods

### 2019-12-27 NAR_form_article.Rmd
NAR_submission files 27 Dec 2019

---
title: Network based multifactorial modelling microRNA-target interaction.
author:
  - name: Selcen Ari Yuka
    affiliation: "1"
  - name: Alper Yilmaz
    affiliation: "1"
    correspond: "To whom correspondence should be addressed. Phone: +90 212 383 4627 Fax: +90 212 383 4625  Email: alyilmaz@yildiz.edu.tr"
    separator: "and"
address:
  - code: 1
    address: Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey

abstract: |
  Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is remarkable in means of miRNA regulation. Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions. In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values. Our model is able to handle miRNA interaction factors such as seed type, binding energy, if provided. Our approach is able to reveal that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be calculated. Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies. We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs. We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods. A gene over-expressed in tumor tissue, namely *SERPINE2*, is used as perturbing element. We have observed that change in expression level of single gene in miRNA:target network is sufficient to perturb regulations in whole network, due to unforeseen and unpredicted regulation which are only visible when considered in network context. Therefore, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account. The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs. The model is available as R package ceRNAnetsim (https://github.com/selcenari/ceRNAnetsim).

bibliography: NAR-sample.bib
output: rticles::nar_article
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \usepackage[colorlinks=true,
            linkcolor=blue,
            urlcolor=blue,
            citecolor=blue]{hyperref}
  \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

As it is possible for miRNAs to suppress multiple targets, an individual mRNA molecule can also be targeted by multiple miRNAs. In that case, the targeted mRNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs. Briefly, Ala et al. have explained the ceRNA hypothesis as disturbance of the other target when one of the targets on a steady-state system that included one miRNA and two target was perturbed with expression change [@ala_integrated_2013]. Regarding interaction between miRNAs and their target in a cell, explaining and predicting results of an individual perturbation is difficult due to complexity of interactions. Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions.

@@

@@

# Genes causing widespread perturbation

Additionally, miRNA:gene expression dataset which was integrated via miRTarBase [@chou2017mirtarbase] was analysed in mean of perturbation efficiency to determine crucial genes on the network. We have observed that one miRNA and 297 genes have high perturbation efficiency, when we ran our model with more than 17% efficiency on target expression. The miRNA, miR-10b was investigated via HMDD database [@huang2018hmdd] and was found as metastatic biomarker. The genes with high perturbation efficiency were analysed via eDGAR, a database of Disease-Gene Associations with annotated relationships among genes [@babbi2017edgar] that supplies different observations about crucial genes such as direct physical interactions, other shared genes with these interactions, protein-protein interactions, GO terms etc.  With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast cancer genes and these genes show direct physical interaction with 21 genes (Figure S10). 33 shared genes which are observed on these direct interactions were also found in our effective perturbation elements. We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset (Figure S10). 44 of regulated genes via these TFs were found in our highly effective genes. it has been observed that these genes cover 22.5% of 297 perturbation-effective genes.

When we analysed these genes via biomaRt OMIM query, it has been observed that our model provides enriched 1.45 fold results with hypergeometric p-values 3.19e-05.

>>>>

---
title: Network based multifactorial modelling microRNA-target interaction.
author:
  - name: Selcen Ari
    affiliation: "1"
  - name: Alper Yilmaz
    affiliation: "1"
    correspond: "To whom correspondence should be addressed. Phone: +90 212 383 4627 Fax: +90 212 383 4625  Email: alyilmaz@yildiz.edu.tr"
    separator: "and"
address:
  - code: 1
    address: Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey

abstract: |
  Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is an important and under-explored subject. Several studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions. In this paper, we defined a new network-based model that incorporates miRNA:ceRNA interactions with expression values. Our approach is able to calculate network-wide effects after a perturbation in expression level of a single element while utilizing miRNA interaction factors such as seed type, binding energy. We applied the model to existing breast cancer mRNA, miRNA expression and high throughput experimental miRNA:target interaction dataset. A gene over-expressed in tumor tissue, namely *SERPINE2*, was identified as a perturbing element which affect regulation of many nodes in whole network when its expression level changes. To conclude, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account. The model has potential to reveal unforeseen and unpredicted regulation which are only visible when considered in network context. Our tool is scalable and can be plugged in with emerging miRNA effectors such as circRNAs and available as R package ceRNAnetsim (https://github.com/selcenari/ceRNAnetsim)

bibliography: NAR-sample
output: rticles::nar_article
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \usepackage[colorlinks=true,
            linkcolor=blue,
            urlcolor=blue,
            citecolor=blue]{hyperref}
  \usepackage[T1]{fontenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)
```

@@

As it is possible for miRNAs to suppress multiple targets, an individual mRNA molecule can also be targeted by multiple miRNAs. In that case, the targeted mRNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs. Briefly, Ala et al. have explained the ceRNA hypothesis as disturbance of the other target when one of the targets on a steady-state system that included one miRNA

@@

and two target was perturbed with expression change [@ala_integrated_2013]. Regarding interaction between miRNAs and their target in a cell, explaining and predicting results of an individual perturbation is difficult due to complexity of interactions. Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions.

@@

## Genes causing widespread perturbation

Additionally, miRNA:gene expression dataset which was integrated via miRTarBase [@chou2017mirtarbase] was analysed in mean of perturbation efficiency to determine crucial genes on the network. We have observed that one miRNA and 297 genes have high perturbation efficiency, when we ran our model with more than 17% efficiency on target expression. The miRNA, miR-10b was investigated via HMDD database [@huang2018hmdd] and was found as metastatic biomarker. The genes with high perturbation efficiency were analysed via eDGAR, a database of Disease-Gene Associations with annotated relationships among genes [@babbi2017edgar] that supplies different observations about crucial genes such as direct physical interactions, other shared genes with these interactions, protein-protein interactions, GO terms etc.  With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast cancer genes and these genes show direct physical interaction with 21 genes (Figure S10). 33 shared genes which are observed on these direct interactions were also found in our effective perturbation elements. We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset (Figure S10). 44 of regulated genes via these TFs were found in our highly effective genes. It has been observed that these genes cover 22.5% of 297 perturbation-effective genes. When we analysed these genes via biomaRt OMIM query, they show 1.45 fold enrichment of OMIM genes (hypergeometric p-value 3.19e-05).

### 2019-12-27 NAR_form_article.tex
NAR_submission files 27 Dec 2019

% Enter dates of publication
\copyrightyear{2008}
\pubdate{31 July 2009}
\pubyear{2009}
\jvolume{37}
\jissue{12}

@@

@@

\author{
    Selcen Ari\,$^{1}$
   and Alper Yilmaz\,$^{1}$ \footnote{To whom correspondence should be addressed. Phone: +90 212 383 4627 Fax:
+90 212 383 4625 Email:
\href{mailto:alyilmaz@yildiz.edu.tr}{\nolinkurl{alyilmaz@yildiz.edu.tr}}} 
}

@@

\history{%
Received January 1, 2009;
Revised February 1, 2009;
Accepted March 1, 2009}

@@

\begin{abstract}
Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is remarkable in means of miRNA
regulation. Many studies have pointed out that an alteration in
miRNA:target interaction can result in unexpected changes due to
indirect and complex interactions. In this paper, we defined a new
network-based model that handles miRNA:ceRNA interactions with
expression values. Our model is able to handle miRNA interaction factors
such as seed type, binding energy, if provided. Our approach is able to
reveal that a perturbation in an element of network affects whole
competing elements differently and cooperative efficiencies of miRNAs on
common targets could be calculated. Our findings emphasized importance
of miRNA:target ratios being crucial, as reported by previous studies.
We have showed that the competing elements which have the same or close
expression values may not be affected equally from the perturbation
because of repression functionality depended on interaction factors of
miRNA target pairs. We applied the model to real sample consisting of
breast cancer gene and miRNA expression dataset and experimental
miRNA:target interaction dataset all generated via high throughput
sequencing methods. A gene over-expressed in tumor tissue, namely
\emph{SERPINE2}, is used as perturbing element. We have observed that
change in expression level of single gene in miRNA:target network is
sufficient to perturb regulations in whole network, due to unforeseen
and unpredicted regulation which are only visible when considered in
network context. Therefore, this model helps unveiling the crosstalk
between elements in miRNA:target network where abundance of target and
sponge effect are taken into account. The model is scalable and can be
plugged in with emerging miRNA effectors such as circRNAs. The model is
available as R package ceRNAnetsim
(\url{https://github.com/selcenari/ceRNAnetsim}).
\end{abstract}

@@

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs)
\citep{ala_integrated_2013, cesana_deciphering_2013}, against their
miRNAs. Briefly, Ala et al.~have explained the ceRNA hypothesis as
disturbance of the other target when one of the targets on a
steady-state system that included one miRNA and two target was perturbed
with expression change \citep{ala_integrated_2013}. Regarding
interaction between miRNAs and their target in a cell, explaining and
predicting results of an individual perturbation is difficult due to
complexity of interactions. Various computational and experimental
studies have tackled the problem of unraveling ceRNA:miRNA interactions.

@@

@@

\subsection{Simulation of large and realistic network}

@@

\section{Additions to be fixed}

Additionally, miRNA:gene expression dataset which was integrated via
miRTarBase \citep{chou2017mirtarbase} was analysed in mean of
perturbation efficiency to determine crucial genes on the network. We
have observed that two miRNAs and 488 genes are effective in terms of
perturbation efficiency, when we ran our model with more than 15\%
efficiency on target expression. miR-10b was investigated via HMDD
database \citep{huang2018hmdd} and mostly found as metastatic biomarker.
On the other hand, it was observed that miR-148a is critical for
promotion of apoptosis and supression of breast cancer cell growth.
Additionaly, to obtain the number of genes in a narrow range, the model
was ran with 17\% efficency on target expression and 297 genes and one
miRNA were observed as highly perturbation-effective. We analysed these
genes via eDGAR, a database of Disease-Gene Associations with annotated
Relationships among genes \citep{babbi2017edgar} that supplies different
observations about crucial genes such as direct physical interactions,
other shared genes with these interactions, protein-protein
interactions, GO terms etc. With our model, we obtained 3 (TP53, BRCA1,
AKT1) of 27 critical breast cancer genes and these genes show direct
physical interaction with 21 genes. 33 shared genes which are observed
on these direct interactions were also found in our effective
perturbation elements.We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4,
GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical
TFs which are provided by eDGAR dataset. 44 of regulated genes via these
TFs were found in our highly effective genes. it has been observed that
these genes cover 22.5\% of 297 perturbation-effective genes.

When we analysed these genes via biomaRt OMIM query, it has been
observed that our model provides enriched 1.45 fold results with
hypergeometric p-values 3.19e-05.

@@

\bibliographystyle{nar}
\bibliography{NAR-sample.bib}

>>>>

% Enter dates of publication
\copyrightyear{X}
\pubdate{X}
\pubyear{X}
\jvolume{X}
\jissue{X}

@@

%\usepackage[scaled]{helvet}
%\renewcommand\familydefault{\sfdefault}

@@

\author{
    Selcen Ari Yuka\,$^{1}$ \footnote{To whom correspondence should be addressed. Email: selcenay@yildiz.edu.tr}
   and Alper Yilmaz\,$^{1}$ \footnote{Correspondence may also be addressed to Wieteke Alper Yilmaz. Phone: +90 212 383 4627 Fax: +90 212 383 4625 Email: alyilmaz@yildiz.edu.tr}
}

@@

\history{%
X;
X;
X}

@@

\begin{abstract}
Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is an important and
under-explored subject. Several studies have pointed out that an
alteration in miRNA:target interaction can result in unexpected changes
due to indirect and complex interactions. In this paper, we defined a
new network-based model that incorporates miRNA:ceRNA interactions with
expression values. Our approach is able to calculate network-wide
effects after a perturbation in expression level of a single element
while utilizing miRNA interaction factors such as seed type, binding
energy. We applied the model to existing breast cancer mRNA, miRNA
expression and high throughput experimental miRNA:target interaction
dataset. A gene over-expressed in tumor tissue, namely \emph{SERPINE2},
was identified as a perturbing element which affect regulation of many
nodes in whole network when its expression level changes. To conclude,
this model helps unveiling the crosstalk between elements in
miRNA:target network where abundance of target and sponge effect are
taken into account. The model has potential to reveal unforeseen and
unpredicted regulation which are only visible when considered in network
context. Our tool is scalable and can be plugged in with emerging miRNA
effectors such as circRNAs and available as R package ceRNAnetsim
(\url{https://github.com/selcenari/ceRNAnetsim})
\end{abstract}

@@

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs)
\citep{ala_integrated_2013, cesana_deciphering_2013}, against their
miRNAs. Briefly, Ala et al.~have explained the ceRNA hypothesis as
disturbance of the other target when one of the targets on a
steady-state system that included one miRNA

@@

and two target was perturbed with expression change
\citep{ala_integrated_2013}. Regarding interaction between miRNAs and
their target in a cell, explaining and predicting results of an
individual perturbation is difficult due to complexity of interactions.
Various computational and experimental studies have tackled the problem
of unraveling ceRNA:miRNA interactions.

@@

\subsection{Simulation of a large and realistic network}

@@

\subsection{Genes causing widespread perturbation}

Additionally, miRNA:gene expression dataset which was integrated via
miRTarBase \citep{chou2017mirtarbase} was analysed in mean of
perturbation efficiency to determine crucial genes on the network. We
have observed that one miRNA and 297 genes have high perturbation
efficiency, when we ran our model with more than 17\% efficiency on
target expression. The miRNA, miR-10b was investigated via HMDD database
\citep{huang2018hmdd} and was found as metastatic biomarker. The genes
with high perturbation efficiency were analysed via eDGAR, a database of
Disease-Gene Associations with annotated relationships among genes
\citep{babbi2017edgar} that supplies different observations about
crucial genes such as direct physical interactions, other shared genes
with these interactions, protein-protein interactions, GO terms etc.
With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast
cancer genes and these genes show direct physical interaction with 21
genes (Figure S10). 33 shared genes which are observed on these direct
interactions were also found in our effective perturbation elements. We
have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1)
transcription factors (TFs) that are common in 44 critical TFs which are
provided by eDGAR dataset (Figure S10). 44 of regulated genes via these
TFs were found in our highly effective genes. It has been observed that
these genes cover 22.5\% of 297 perturbation-effective genes.When we 
analysed these genes via biomaRt OMIM query, they show 1.45 fold 
enrichment of OMIM genes (hypergeometric p-value 3.19e-05).

@@

\bibliographystyle{NAR-natbib}
\bibliography{NAR-sample}

### 2019-12-27 Supplementary_additional_file.Rmd
NAR_submission files 27 Dec 2019

---
title: "Supplementary Experimental Data File"
author: 
- Selcen Ari
- Alper Yilmaz 
date: "22 09 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: ref.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

@@

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(readxl)
library(biomaRt)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)

@@

```{r,  eval=FALSE}
#HG19
listEnsemblArchives()
listMarts(host = 'http://grch37.ensembl.org')
ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')
hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)

@@

```{r, eval=FALSE}
read_tsv("mirna_mature.txt", col_names = FALSE)%>%
  filter(startsWith(X2, "hsa"))%>%
  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
  inner_join(mirbasehg38%>% dplyr::select(ID, Name), by= c("mirbase_ID"= "ID"))%>%
  dplyr::select(mirbase_ID, Name)%>%
  distinct()%>%
  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, seed_type, seed_basepairs,folding_class, seq_ID, folding_energy, X5.UTR, CDS, X3.UTR)%>%
  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", "Ensembl_Transcript_Id"="ensembl_transcript_id", "Hugo_Symbol"="hgnc_symbol"))%>%
  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), region2= ifelse(X3.UTR == "1", "3UTR", " "),region3= ifelse(CDS =="1", "CDS", " "))%>%
  unite(region, c(region1, region2, region3), sep = "||")%>%
  dplyr::select(chromosome_name, start_position, end_position,strand, Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type,seed_basepairs, folding_class, folding_energy, region)%>%
  as_tibble() -> clashelwakfinal

@@

hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
```

@@

Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study. On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study. Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values. We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]. With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.

>>>>

---
title: "Supplementary Experimental Data File"
author: 
- Selcen Ari Yuka
- Alper Yilmaz 
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default#
bibliography: NAR-sample.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

@@

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(readxl)
library(biomaRt)

@@

```{r,  eval=FALSE}
#HG19
listEnsemblArchives()
listMarts(host = 'http://grch37.ensembl.org')
ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')
hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id",
                             "chromosome_name", "start_position","end_position", 
                             "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)

@@

```{r, eval=FALSE}
read_tsv("mirna_mature.txt", col_names = FALSE)%>%
  filter(startsWith(X2, "hsa"))%>%
  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
  inner_join(mirbasehg38%>% dplyr::select(ID, Name), 
             by= c("mirbase_ID"= "ID"))%>%
  dplyr::select(mirbase_ID, Name)%>%
  distinct()%>%
  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, 
                Ensembl_Transcript_Id, Hugo_Symbol, 
                mRNA_seq_extended, chimeras_decompressed, 
                seed_type, seed_basepairs,folding_class, 
                seq_ID, folding_energy, 
                X5.UTR, CDS, X3.UTR)%>%
  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", 
                          "Ensembl_Transcript_Id"="ensembl_transcript_id", 
                          "Hugo_Symbol"="hgnc_symbol"))%>%
  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), 
         region2= ifelse(X3.UTR == "1", "3UTR", " "),
         region3= ifelse(CDS =="1", "CDS", " "))%>%
  unite(region, c(region1, region2, region3), sep = "||")%>%
  dplyr::select(chromosome_name, start_position, 
                end_position, strand, Hugo_Symbol, 
                Ensembl_Gene_Id, Ensembl_Transcript_Id, 
                mRNA_seq_extended, Name, miRNA_seq, 
                seq_ID, seed_type,seed_basepairs, 
                folding_class, folding_energy, region)%>%
  as_tibble() -> clashelwakfinal

@@

hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id",
                             "chromosome_name", "start_position","end_position", 
                             "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
```

@@

Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study. On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.Grimson et al. [-@grimson_microrna_2007] have compared the effect of seed types on target repression with few miRNA had canonical seed pairing in their study. Additionally, Bartel et al. [-@bartel_micrornas:_2009] and Betel et al. [-@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values. We also have redefined location of binding region on the target as numeric values, based on studies of Hausser et al. [-@hausser_analysis_2013] and Helwak et al. [-@helwak_mapping_2013]. With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.

### 2019-12-27 Supplementary_figures.Rmd
NAR_submission files 27 Dec 2019

---
title: "Supplementary Figures"
author: 
- Selcen Ari
- Alper Yilmaz 
date: "17 09 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: ref.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

@@

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)

@@

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4",
            title = "Minimal dataset in steady-state conditions")
```

```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
            title = "Gene2 Upregulation without interaction factors")
```

@@

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", 
            title = "Minimal dataset with interaction factors in steady-state conditions")

@@

```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
            title = "Gene2 Upregulation with interaction factors")
```

@@

## Determination of iteration for Upregulation of SERPINE2 gene from Breast cancer patient dataset

>>>>

---
title: "Supplementary Figures"
author: 
- Selcen Ari Yuka
- Alper Yilmaz 
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

@@

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)

@@

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4",
            title = "Minimal dataset in\nsteady-state conditions")
```

```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "Gene2 Upregulation\nwithout interaction factors")
```

@@

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            title = "Minimal dataset with interaction\nfactors in steady-state conditions")

@@

```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "Gene2 Upregulation\nwith interaction factors")
```

@@

## Determination of optimum iteration for Upregulation of SERPINE2 gene from Breast cancer patient dataset

### 2019-12-27 Supplementary_mat_met.Rmd
NAR_submission files 27 Dec 2019

---
title: "Supplementary Materials And Methods"
author: 
- Selcen Ari
- Alper Yilmaz 
date: "7 Nisan 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: ref.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

@@

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)

@@

There are various datasets about miRNA:target pairs such as miRTarBase, DianaTools, miRecords, miRWalk etc. Some of these present the experimentally supported miRNA target pairs or only predicted ones. The experimentally supported datasets generally provides weak evidence for interactions. For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al. and Moore et al. These steps were not handle in this file because they contain many processes.

@@

>>>>

---
title: "Supplementary Materials And Methods"
author: 
- Selcen Ari Yuka
- Alper Yilmaz 
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: NAR-sample.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

@@

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
#knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)

@@

There are various datasets about miRNA:target pairs such as miRTarBase, DianaTools, miRecords, miRWalk etc. Some of these present the experimentally supported miRNA target pairs or only predicted ones. The experimentally supported datasets generally provides weak evidence for interactions. For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al. [-@helwak_mapping_2013] and Moore et al. [-@moore_mirnatarget_2015] These steps were not handle in this file because they contain many processes.

@@

# REFERENCES
